---
title: "PRO-seq Analysis Report"
author: "nf_PROseq Pipeline"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
params:
  results_dir: "."
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
suppressWarnings(suppressMessages({
    library(ggplot2)
    library(knitr)
    library(openxlsx)
}))

results_dir <- params$results_dir
```

# Summary

This report summarizes the results of the PRO-seq (Precision Run-On sequencing) analysis pipeline.

**Pipeline:** nf_PROseq (Nextflow DSL2)
**Date:** `r Sys.Date()`
**Results directory:** `r results_dir`

```{r overview}
# List all output files
all_files <- list.files(results_dir, recursive = TRUE, full.names = FALSE)
cat(sprintf("- %s\n", head(all_files, 50)))
if (length(all_files) > 50) {
    cat(sprintf("  ... and %d more files\n", length(all_files) - 50))
}
```

# Quality Control

## Read Statistics

```{r qc-stats, fig.width=8, fig.height=6}
qc_dir <- file.path(results_dir, "03.Data_QC")
stat_file <- file.path(qc_dir, "read_statistics.txt")

if (file.exists(stat_file)) {
    stats <- read.delim(stat_file, header = TRUE, stringsAsFactors = FALSE)
    kable(stats, caption = "Read QC Statistics")
} else {
    cat("*QC statistics file not found.*")
}
```

## Alignment Summary

```{r alignment}
align_dir <- file.path(results_dir, "04.Alignment")
log_files <- list.files(align_dir, pattern = "summary_bowtie2", full.names = TRUE, recursive = TRUE)

if (length(log_files) > 0) {
    for (lf in log_files) {
        cat(sprintf("\n### %s\n\n", basename(lf)))
        lines <- readLines(lf)
        cat(paste(lines, collapse = "\n"))
        cat("\n")
    }
} else {
    cat("*Alignment logs not found.*")
}
```

# Pausing Index

```{r pausing-index}
pi_dir <- file.path(results_dir, "06.Pausing_Index")
pi_file <- file.path(pi_dir, "Pausing_Index_All.xlsx")

if (file.exists(pi_file)) {
    pi_data <- read.xlsx(pi_file, sheet = 1)
    cat(sprintf("**Total genes with pausing index:** %d\n\n", nrow(pi_data)))

    pi_cols <- grep("_PI$", names(pi_data), value = TRUE)
    if (length(pi_cols) > 0) {
        cat("\n### Pausing Index Summary\n\n")
        for (pc in pi_cols) {
            vals <- pi_data[[pc]]
            vals <- vals[is.finite(vals) & vals > 0]
            cat(sprintf("- **%s**: median = %.3f, mean = %.3f, n = %d\n",
                       pc, median(vals, na.rm = TRUE),
                       mean(vals, na.rm = TRUE), length(vals)))
        }
    }
} else {
    cat("*Pausing index results not found.*")
}
```

# Differential Expression

```{r de}
de_dir <- file.path(results_dir, "07.Diff")
de_file <- file.path(de_dir, "Diff_genes.xlsx")

if (file.exists(de_file)) {
    de_data <- read.xlsx(de_file, sheet = 1)

    if (nrow(de_data) > 0 && !all(is.na(de_data$gene_id))) {
        comparisons <- unique(de_data$Comparison)
        cat(sprintf("**Comparisons:** %s\n\n", paste(comparisons, collapse = ", ")))

        for (comp in comparisons) {
            comp_data <- de_data[de_data$Comparison == comp, ]
            n_up   <- sum(comp_data$Significant == "Up")
            n_down <- sum(comp_data$Significant == "Down")
            cat(sprintf("\n### %s\n\n", comp))
            cat(sprintf("- Up-regulated: %d\n", n_up))
            cat(sprintf("- Down-regulated: %d\n", n_down))
            cat(sprintf("- Total: %d\n\n", n_up + n_down))

            # Top 10 genes
            if (nrow(comp_data) > 0) {
                top <- head(comp_data[order(comp_data$FDR), ], 10)
                kable(top[, c("gene_id", "logFC", "FDR", "Significant")],
                      caption = paste("Top 10 DEGs:", comp))
            }
        }
    } else {
        cat("*No differentially expressed genes found.*")
    }
} else {
    cat("*Differential expression results not found.*")
}
```

# Enrichment Analysis

```{r enrichment}
go_dir  <- file.path(results_dir, "08.GO_KEGG", "GO_results")
kegg_dir <- file.path(results_dir, "08.GO_KEGG", "KEGG_results")
gsea_dir <- file.path(results_dir, "09.GSEA", "GSEA_results")

cat(sprintf("- GO results: %s (exists: %s)\n", go_dir, dir.exists(go_dir)))
cat(sprintf("- KEGG results: %s (exists: %s)\n", kegg_dir, dir.exists(kegg_dir)))
cat(sprintf("- GSEA results: %s (exists: %s)\n", gsea_dir, dir.exists(gsea_dir)))
```

# Figures

## PCA Plot

```{r pca-fig, fig.width=8, fig.height=7}
pca_file <- file.path(results_dir, "10.Plot", "PCA_Plot.pdf")
cat(sprintf("PCA plot: %s (exists: %s)\n", pca_file, file.exists(pca_file)))
```

## Volcano Plot

```{r volcano-fig, fig.width=8, fig.height=7}
volcano_file <- file.path(results_dir, "10.Plot", "Volcano_Plot.pdf")
cat(sprintf("Volcano plot: %s (exists: %s)\n", volcano_file, file.exists(volcano_file)))
```

## Heatmap

```{r heatmap-fig, fig.width=10, fig.height=8}
heatmap_file <- file.path(results_dir, "10.Plot", "Heatmap.pdf")
cat(sprintf("Heatmap: %s (exists: %s)\n", heatmap_file, file.exists(heatmap_file)))
```

## Pausing Index Boxplot

```{r pi-boxplot, fig.width=8, fig.height=6}
boxplot_file <- file.path(results_dir, "06.Pausing_Index", "Pausing_Index_Boxplot.pdf")
cat(sprintf("Pausing Index Boxplot: %s (exists: %s)\n", boxplot_file, file.exists(boxplot_file)))
```

# Metagene Analysis

```{r metagene}
meta_dir <- file.path(results_dir, "11.Metagene")
meta_files <- list.files(meta_dir, pattern = "\\.pdf$", recursive = TRUE)
cat("Metagene plots:\n")
for (mf in meta_files) {
    cat(sprintf("- %s\n", mf))
}
```

# Output Files

```{r output-list}
cat("Complete output file list:\n\n")
for (f in all_files) {
    cat(sprintf("- `%s`\n", f))
}
```
